---
title: "Farm dam emissions & NP data report"
author: "Dr Pawel Waryszak"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  word_document: default
  html_document:
    keep_md: true
    code_folding: hide
    warning: false
    message: false
    collapse: true
---


```{r setup, include=FALSE, message=FALSE, warning=FALSE}

# Load necessary libraries ( e.g., install.packages("knitr") if not present on your comp)
library(tidyverse)
library(dplyr)
library(lubridate)
library(readxl)
library(knitr)
library(gt)

#SET WORKING DIRECTORY where data excel files are:
setwd("C:/Users/poles/Documents/00Deakin_Docs/R/BCL_R/FarmDams/DATA_Toowoomba")

# Define MyFarmer for this report
MyFarmer <- "Karen_Jarling"

# Read the data
field_data <- read_excel("Farm_Dams_Toowoomba_Data.xlsx", sheet = "FarmDams") #continuously updated on TEAMs (see TealCarbonGroup)
water_data <- read_excel("Farm_Dams_Toowoomba_Data.xlsx", sheet = "WATER_DATA") #continuously updated on TEAMs (see TealCarbonGroup)

emission_data <- read.csv("RAW_MethaneData_Karen_Jarling_All.csv") #also on Teams downloaded from deployed Pondi.
emission_data$Date.Time.UTC.10 <-   as.POSIXct(emission_data$Date.Time.UTC.10 ,"%d/%m/%Y %H:%M",  tz="UTC") #convert to time format that R works with


# Filter field_data for MyFarmer
farms <- field_data %>% 
  filter(grepl(MyFarmer, FarmDam_ID)) %>% 
  select(FarmDam_ID, TimeStart, TimeEnd)



```


## Introduction

This document reports the average methane (CH4)and carbon dioxide (CO2) emissions (+-Standard error bars) as recorded by Pondi - our floating sensor. The water nitrogen (N) and phosphorus (P) content are preseted as reported by Eurofins lab in Brisbane based off ~1L water samples we collected  during the farm dam campaign in Queensland at the farm owned by `r MyFarmer` where we tested the following farm dams:

```{r}

# Create a nice table with gt
farms %>%
  gt() %>%
  tab_header(
    title = "Farm Dam Information",
    subtitle = "Data for Selected Farmer"
  ) %>%
  cols_label(
    FarmDam_ID = "Farm Dam ID",
    TimeStart = "Start Time",
    TimeEnd = "End Time"
  ) %>%
  tab_style(
    style = list(
      cell_fill(color = "lightblue"),
      cell_text(weight = "bold")
    ),
    locations = cells_title(groups = "title")
  ) %>%
  tab_style(
    style = cell_borders(sides = "all", weight = px(1)),
    locations = cells_body()
  )


```


## CO2 emission

```{r, include=TRUE, message=FALSE, warning=FALSE}

# Define a function to process and plot data for a given FarmDam_ID
plot_emissions <- function(farmdam_id) {
  # Extract start and end times for the current farm dam
  start_time <- farms %>% 
    filter(FarmDam_ID == farmdam_id) %>% 
    select(TimeStart) %>% 
    pull()
  
  end_time <- farms %>% 
    filter(FarmDam_ID == farmdam_id) %>% 
    select(TimeEnd) %>% 
    pull()
  
  emission_start_time <- as.POSIXct(strptime(start_time, "%d/%m/%Y %H:%M", tz="UTC"))
  emission_end_time <- as.POSIXct(strptime(end_time, "%d/%m/%Y %H:%M", tz="UTC"))
  
  start <- as.Date(emission_start_time)
  end <- as.Date(emission_end_time)
  
  # Filter and summarize CO2 data
  CO2_filtered_data <- emission_data %>% 
    filter(Date.Time.UTC.10 >= emission_start_time & Date.Time.UTC.10 <= emission_end_time) %>% 
    filter(Total.CO2.Est...mg.m.2.day. > 0) %>% # Remove negative emission logs
    group_by(Site) %>%
    summarise(Latitude = mean(Latitude..DD., na.rm = TRUE),
              Longitude = mean(Longitude..DD., na.rm = TRUE),
              AV = mean(Total.CO2.Est...mg.m.2.day. , na.rm = TRUE),
              SD = sd(Total.CO2.Est...mg.m.2.day.) ,
              N = length(Total.CO2.Est...mg.m.2.day.) ,
              SE = SD / sqrt(N)) %>%
    mutate(Gas = "CO2")
  
  # Generate a barplot for CO2 emissions
  plot <- ggplot(CO2_filtered_data, aes(x = Site, y = AV)) +
    geom_bar(position = "identity", stat = "identity", fill = "#b2df8a") +
    geom_errorbar(aes(ymin = AV - SE, ymax = AV + SE), width = 0.8, size = 1.1) +
    
      geom_text(aes(label = paste0("n = ", N)), 
            position = position_stack(vjust = 0.5), 
            vjust = -0.5, 
            size = 6) +
    
    labs(title = paste("Farm Dam ID:", farmdam_id, "[", start, " to ", end, "]"),
         subtitle = paste("Located at:", CO2_filtered_data$Latitude[1], ",", CO2_filtered_data$Longitude[1]),
         x = "", y = bquote("CO"[2] ~ "flux " ~(mg * ~m^-2 ~day^-1))) +
    theme_bw() +
    theme(axis.text.y = element_text(size = 10),
          axis.title.x = element_text(size = 14),
          axis.text.x = element_text(size = 14),
          axis.title.y = element_text(size = 14),
          strip.text = element_text(size = 16),
          legend.position = "right")
  
  # Print the plot
  print(plot)
}

# Apply the function to each farm dam ID
for (farmdam_id in farms$FarmDam_ID) {
  plot_emissions(farmdam_id)
}
```



## CH4 emission

```{r, include=TRUE, message=FALSE, warning=FALSE}

# Define a function to process and plot data for a given FarmDam_ID
plot_emissions2 <- function(farmdam_id) {
  # Extract start and end times for the current farm dam
  start_time <- farms %>% 
    filter(FarmDam_ID == farmdam_id) %>% 
    select(TimeStart) %>% 
    pull()
  
  end_time <- farms %>% 
    filter(FarmDam_ID == farmdam_id) %>% 
    select(TimeEnd) %>% 
    pull()
  
  emission_start_time <- as.POSIXct(strptime(start_time, "%d/%m/%Y %H:%M", tz="UTC"))
  emission_end_time <- as.POSIXct(strptime(end_time, "%d/%m/%Y %H:%M", tz="UTC"))
  
  start <- as.Date(emission_start_time)
  end <- as.Date(emission_end_time)
  
  # Filter and summarize CO2 data
  CH4_filtered_data <- emission_data %>% 
    filter(Date.Time.UTC.10 >= emission_start_time & Date.Time.UTC.10 <= emission_end_time) %>% 
    filter(Total.Methane.Est...mg.m.2.day. > 0) %>% # Remove negative emission logs
    group_by(Site) %>%
    summarise(Latitude = mean(Latitude..DD., na.rm = TRUE),
              Longitude = mean(Longitude..DD., na.rm = TRUE),
              AV = mean(Total.Methane.Est...mg.m.2.day. , na.rm = TRUE),
              SD = sd(Total.Methane.Est...mg.m.2.day. ),
              N = length(Total.Methane.Est...mg.m.2.day. ),
              SE = SD / sqrt(N)) %>%
    mutate(Gas = "CH4")
  
  # Generate a barplot for emissions
  plot <- ggplot(CH4_filtered_data, aes(x = Site, y = AV)) +
    geom_bar(position = "identity", stat = "identity", fill = "#1f78b4") +
    geom_errorbar(aes(ymin = AV - SE, ymax = AV + SE), width = 0.8, size = 1.1) +
    
      geom_text(aes(label = paste0("n = ", N)), 
            position = position_stack(vjust = 0.5), 
            vjust = -0.5, 
            size = 6) +
    
    labs(title = paste("Farm Dam ID:", farmdam_id, "[", start, " to ", end, "]"),
         subtitle = paste("Located at:", CH4_filtered_data$Latitude[1], ",", CH4_filtered_data$Longitude[1]),
         x = "", y = bquote("CH"[4] ~ "flux " ~(mg * ~m^-2 ~day^-1))) +
    theme_bw() +
    theme(axis.text.y = element_text(size = 10),
          axis.title.x = element_text(size = 14),
          axis.text.x = element_text(size = 14),
          axis.title.y = element_text(size = 14),
          strip.text = element_text(size = 16),
          legend.position = "right")
  
  # Print the plot
  print(plot)
}

# Apply the function to each farm dam ID
for (farmdam_id in farms$FarmDam_ID) {
  plot_emissions2(farmdam_id)
}
```

## Water Testing Results

```{r, message=FALSE, warning=FALSE}
# Subset columns containing "KAREN_JARLING"
karen_jarling_data <- water_data %>%
  select(contains(MyFarmer) | contains("sample"))

# Select rows 7 and 9 (Total N and P)
selected_rows <- karen_jarling_data [c(7, 9), ] #select NP total rows

gathered_rows <- gather(selected_rows,key = "FarmDam", value = "Value",-Client_SampleID) %>%
  rename(NP = Client_SampleID ) %>% 
  select(FarmDam, NP, Value) # reorder the columns



# Create a nice table with gt
gathered_rows  %>%
  gt() %>%
  tab_header(
    title = "Farm Dam Water Quality",
    subtitle = "Nitrogen (N)  and Phosphate (P) content"
  ) %>%
  cols_label(
    FarmDam = "Farm Dam ID",
     NP = "Total NP",
    Value = "Value"
  ) %>%
  tab_style(
    style = list(
      cell_fill(color = "lightblue"),
      cell_text(weight = "bold")
    ),
    locations = cells_title(groups = "title")
  ) %>%
  tab_style(
    style = cell_borders(sides = "all", weight = px(1)),
    locations = cells_body()
  )
```

